---
title: "Analysis and visualisations"
output: html_notebook
---
# Load libraries ----
```{r}
library(tidyverse)
library(here)
```
# Read in the data ----
```{r}
shs_data <- read_csv(here::here("clean_data/shs_data.csv"))

# Additional data cleaning
shs_data <- shs_data %>% 
  mutate(across(where(is_character), str_to_sentence)) %>% 
  mutate(feature_name = str_to_title(feature_name)) %>% 
  mutate(measure = str_replace_all(measure, "_", " ")) %>% 
  rename(year = date_code)


glimpse(shs_data)
```

# Explore the data ----
(would be nice to do some hypothesis testing)
```{r}
skimr::skim(shs_data)
```

The data is aggregated from several datasets, covering 2013-2019 (`year`). 
The `percent` column is an estimate of the percentage of respondents that answered the survey. 
The `measure` column represents what factor is being measured. 
`response` relates to the response to a particular question (aka `measure`) from the SHS. 
`feature_name`, `feature_code` and `feature_type` all correspond to the location of the respondent. `urban_rural_classification` classifiers the area of a respondent as either being rural or urban. 
`gender` can either be "Male" or "Female". 
`household_type` refers to who lives in the property, "Adults", "Pensioners" or "With children". 
`type_of_tenure` captures whether the property is "Owned", "Owned mortgage/loan", "Private rented", "Social rented" or "Other". 
`SIMD_quintiles` includes the top and lower 20% of the deprivation scale. 
`ethnicity` records the ethnicity of respondents as either "White" or "Other".
`walking_distance_to_nearest_greenspace` refers to the distance respondents have to walk to greenspace. NAs are present where data in the `measure` column relates to distance to greenspace.  
Data is aggregated across all columns to include "All" levels. 


# Visualisations ----

### Are there big differences in how far people have to walk to access their greenspaces?

```{r}
# Across Scotland
distance_to_greenspace_scotland <- shs_data %>% 
  filter(measure == "Distance to nearest green or blue space" &
           feature_name == "Scotland" &
           gender == "All" &
           urban_rural_classification == "All" &
           simd_quintiles == "All" &
           type_of_tenure  == "All" &
           household_type  == "All" &
           ethnicity == "All") %>% 
  group_by(response) %>% 
  mutate(percentage_2013_2019 = mean(percent)) %>% 
  unique() 

distance_scotland_plot <- distance_to_greenspace_scotland %>% 
  distinct(percentage_2013_2019) %>% 
  mutate(response = factor(response, 
                           ordered = TRUE, 
                           levels = c("A 5 minute walk or less",
                                      "Within a 6-10 minute walk",
                                      "An 11 minute walk or more",
                                      "Don't know"))) %>% 
  ggplot(aes(x = response, y = percentage_2013_2019)) +
  geom_col(fill = "deepskyblue4") +
  theme_minimal() +
  labs(title = "Distance to nearest greenspace",
       subtitle = "Scotland 2013-2019") +
  xlab("Distance") +
  ylab("Percentage") +
  coord_flip()

```


```{r}
# Distance by area
distance_by_area <- shs_data %>% 
  filter(measure == "Distance to nearest green or blue space" &
           feature_name != "Scotland" &
           gender == "All" &
           urban_rural_classification == "All" &
           simd_quintiles == "All" &
           type_of_tenure  == "All" &
           household_type  == "All" &
           ethnicity == "All") %>% 
  group_by(response, feature_name) %>% 
  mutate(percentage_2013_2019 = mean(percent)) %>% 
  unique() 

distance_5_min_by_area <- distance_by_area %>% 
  filter(response == "A 5 minute walk or less") %>% 
  distinct(percentage_2013_2019) 
# mutate(percentage_threshold = case_when(
#   percentage_2013_2019 >= 80 ~ "> 80%",
#   percentage_2013_2019 >= 75 & percentage_2013_2019 < 80 ~ "75-80%",
#   percentage_2013_2019 >= 70 & percentage_2013_2019 < 75 ~ "70-75%",
#   percentage_2013_2019 >= 65 & percentage_2013_2019 < 70 ~ "65-70%",
#   percentage_2013_2019 >= 60 & percentage_2013_2019 < 65 ~ "60-65%",
#   percentage_2013_2019 >= 50 & percentage_2013_2019 < 60 ~ "50-60%",
#   percentage_2013_2019 >= 40 & percentage_2013_2019 < 50 ~ "40-50%",
#   TRUE ~ "Other")) %>% 
# mutate(percentage_threshold = factor(percentage_threshold, 
#                                      ordered = TRUE, 
#                                      levels = c("> 80%",
#                                                 "75-80%",
#                                                 "70-75%",
#                                                 "65-70%",
#                                                 "60-65%",
#                                                 "50-60%",
#                                                 "40-50%")))

distance_5_min_by_area$feature_name = with(distance_5_min_by_area, reorder(feature_name, percentage_2013_2019, decreasing = FALSE))

distance_by_area_plot <- distance_5_min_by_area %>% 
  ggplot(aes(x = feature_name, y = percentage_2013_2019)) +
  geom_col(fill = "deepskyblue4") +
  theme_minimal() +
  coord_flip() +
  labs(title = "Percentage of respondents living 5 minutes walk from greenspace",
       subtitle = "2013-2019",
       x = "Area",
       y = "Percentage")

distance_area_plot_2 <- distance_5_min_by_area %>% 
  ggplot(aes(x = feature_name, y = percentage_2013_2019)) +
  geom_segment(aes(x = feature_name, xend = feature_name, y = 0, yend = percentage_2013_2019)) +
  geom_point(size = 1, colour = "deepskyblue4", fill = alpha("deepskyblue4", 0.3), alpha = 0.7, shape = 21, stroke=2) +
  theme_minimal() +
  coord_flip() +
  labs(title = "Percentage of respondents living 5 minutes walk from greenspace",
       subtitle = "Scotland 2013-2019",
       x = "Area",
       y = "Percentage")
```



```{r}
# Distance in neighbourhood rating by distance to greenspace
neighbourhood_rating_distance <- shs_data %>% 
  filter(walking_distance_to_nearest_greenspace %in% c("More than 10 minutes", 
                                                       "Less than 10 minutes")) %>% 
  filter(measure == "Neighbourhood rating") %>% 
  filter(feature_name == "Scotland" &
           gender == "All" &
           urban_rural_classification == "All" &
           simd_quintiles == "All" &
           type_of_tenure  == "All" &
           household_type  == "All" &
           ethnicity == "All" &
           response != "No opinion") %>% 
  group_by(response, walking_distance_to_nearest_greenspace) %>% 
  mutate(percentage_2013_2019 = mean(percent))

neighbourhood_distance_plot <- neighbourhood_rating_distance %>% 
  mutate(response = factor(response, 
                           ordered = TRUE,
                           levels = c("Very good",
                                      "Fairly good", 
                                      "Fairly poor",
                                      "Very poor"))) %>%
  mutate(walking_distance_to_nearest_greenspace = factor(walking_distance_to_nearest_greenspace)) %>% 
  ggplot(aes(x = response, y = percentage_2013_2019, 
             fill = walking_distance_to_nearest_greenspace)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("darkslategray3", "deepskyblue4")) +
  theme_minimal() +
  labs(title = "Neighbourhood rating by walking distance to greenspace",
       subtitle = "Scotland 2013-2019",
       x = "Rating",
       y = "Percentage") +
  guides(fill = guide_legend(title = "Walking distance to nearest greenspace"))


```

```{r}
# Neighbourhood by walking distance urban/rural

# neighbourhood_rating_distance_urban <- shs_data %>% 
#    filter(measure == "Neighbourhood rating") %>% 
#            urban_rural_classification == "Urban" | 
#            urban_rural_classification == "Rural") %>% 
#   filter(response != "No opinion") %>% 
#   group_by(response) %>% 
#   mutate(percentage_2013_2019 = mean(percent))
# 
# neighbourhood_distance_plot_urban <- neighbourhood_rating_distance_urban %>% 
#   mutate(response = factor(response, 
#                            ordered = TRUE,
#                            levels = c("Very good",
#                                       "Fairly good", 
#                                       "Fairly poor",
#                                       "Very poor"))) %>%
#   ggplot(aes(x = response, y = percentage_2013_2019, 
#              fill = urban_rural_classification)) +
#   geom_col(position = "dodge") +
#   scale_fill_manual(values = c("darkslategray3", "deepskyblue4")) +
#   theme_minimal() +
#   labs(title = "Neighbourhood rating by urban or rural classification",
#        subtitle = "Scotland 2013-2019",
#        x = "Rating",
#        y = "Percentage") +
#   guides(fill = guide_legend(title = "Urban or Rural"))
# 
# shs_data %>% 
#   filter(measure == "Neighbourhood rating" & urban_rural_classification == "Urban" | urban_rural_classification == "Rural") %>% 
#   count(walking_distance_to_nearest_greenspace)
```


### Changes over time
```{r}
# Neighbourhood rating
rating_time <- shs_data %>% 
  filter(measure == "Neighbourhood rating") %>% 
  filter(feature_name == "Scotland" &
           gender == "All" &
           urban_rural_classification == "All" &
           simd_quintiles == "All" &
           type_of_tenure  == "All" &
           household_type  == "All" &
           ethnicity == "All" &
           response != "No opinion" &
           walking_distance_to_nearest_greenspace == "All") %>% 
  mutate(response = factor(response, 
                           ordered = TRUE,
                           levels = c("Very good",
                                      "Fairly good", 
                                      "Fairly poor",
                                      "Very poor"))) %>% 
  ggplot(aes(x = year, y = percent, colour = response)) +
  geom_line() +
  scale_color_manual(values = c("deepskyblue4", "darkslategray3", "darkgreen", "darkolivegreen")) +
  theme_minimal() +
  labs(title = "Neighbourhood rating over time",
       subtitle = "Scotland 2013-2019",
       x = "Year",
       y = "Percentage") +
  guides(fill = guide_legend(title = "Rating"))

# over time by area
rating_time_area <- shs_data %>% 
  filter(measure == "Neighbourhood rating") %>% 
  filter(gender == "All" &
           urban_rural_classification == "All" &
           simd_quintiles == "All" &
           type_of_tenure  == "All" &
           household_type  == "All" &
           ethnicity == "All" &
           response != "No opinion" &
           walking_distance_to_nearest_greenspace == "All") %>% 
  mutate(response = factor(response, 
                           ordered = TRUE,
                           levels = c("Very good",
                                      "Fairly good", 
                                      "Fairly poor",
                                      "Very poor"))) %>% 
  ggplot(aes(x = year, y = percent, colour = response)) +
  geom_line() +
  scale_colour_brewer(palette = "BrBG") +
  #scale_color_manual(values = c("deepskyblue4", "darkslategray3", "darkgreen", "darkolivegreen")) +
  theme_minimal() +
  labs(title = "Neighbourhood rating over time",
       subtitle = "Scotland 2013-2019",
       x = "Year",
       y = "Percentage") +
  guides(fill = guide_legend(title = "Rating")) +
  facet_wrap(~feature_name)

rating_time_urban_rural <- shs_data %>% 
  filter(measure == "Neighbourhood rating") %>% 
  filter(feature_name == "Scotland" &
           gender == "All" &
           urban_rural_classification != "All" &
           simd_quintiles == "All" &
           type_of_tenure  == "All" &
           household_type  == "All" &
           ethnicity == "All" &
           response != "No opinion" &
           walking_distance_to_nearest_greenspace == "All") %>% 
  mutate(response = factor(response, 
                           ordered = TRUE,
                           levels = c("Very good",
                                      "Fairly good", 
                                      "Fairly poor",
                                      "Very poor"))) %>% 
  ggplot(aes(x = year, y = percent, colour = response)) +
  geom_line() +
  scale_colour_brewer(palette = "BrBG") +
  theme_minimal() +
  labs(title = "Neighbourhood rating over time",
       subtitle = "Scotland 2013-2019",
       x = "Year",
       y = "Percentage") +
  guides(fill = guide_legend(title = "Rating")) +
  facet_wrap(~ urban_rural_classification)
```

```{r}
shs_data %>% 
  select(feature_name, urban_rural_classification) %>% 
  filter(urban_rural_classification != "All") %>% 
  unique()
```


```{r}
neighbourhood_rating_scotland <- shs_data %>% 
  filter(measure == "Neighbourhood rating" &
           feature_name == "Scotland" &
           gender == "All" &
           urban_rural_classification == "All" &
           simd_quintiles == "All" &
           type_of_tenure  == "All" &
           household_type  == "All" &
           ethnicity == "All") %>% 
  group_by(response) %>% 
  mutate(percentage_2013_2019 = mean(percent)) %>% 
  unique() 


neighbourhood_rating_scotland_plot <- neighbourhood_rating_scotland %>% 
  distinct(percentage_2013_2019) %>% 
  mutate(response = factor(response, 
                           ordered = TRUE,
                           levels = c("Very good",
                                      "Fairly good", 
                                      "Fairly poor",
                                      "Very poor",
                                      "No opinion")))%>% 
  ggplot(aes(x = response, y = percentage_2013_2019)) +
  geom_col(fill = "cyan4") +
  theme_minimal() +
  labs(title = "Neighbourhood rating",
       subtitle = "Scotland 2013-2019") +
  xlab("Rating") +
  ylab("Percentage")
```


## Community Belonging
```{r}
community_belonging_time <- shs_data %>% 
  filter(measure == "Community belonging") %>%
  filter(
    gender == "All" &
      urban_rural_classification == "All" &
      simd_quintiles == "All" &
      type_of_tenure  == "All" &
      household_type  == "All" &
      ethnicity == "All" &
      walking_distance_to_nearest_greenspace == "All") %>% 
  mutate(response = factor(response, 
                           ordered = TRUE,
                           levels = c("Very strongly",
                                      "Fairly strongly", 
                                      "Don't know",
                                      "Not very strongly",
                                      "Not at all strongly"))) %>% 
  ggplot(aes(x = year, y = percent, colour = response)) +
  geom_line() +
  scale_colour_brewer(palette = "BrBG") +
  #scale_color_manual(values = c("deepskyblue4", "darkslategray3", "darkgreen", "darkolivegreen", "darkblue")) +
  facet_wrap(~feature_name) +
  labs(title = "Community belonging over time",
       subtitle = "Scotland 2013-2019",
       x = "Year",
       y = "Percentage") +
  guides(fill = guide_legend(title = "Rating")) +
  theme_minimal()

```

```{r}

```



```{r}
shs_data %>% 
  filter(measure == "Community belonging") %>% 
  distinct(response)
```

# Read in individualised data ----

```{r}
shs_responses <- read_csv(here::here("clean_data/shs_responses_clean.csv"))
```

```{r}
skimr::skim(shs_responses)
shs_responses %>% 
  count(year)
```

```{r}
# neighbourhood rating by community belonging 

neighbourhood_community_rating <- shs_responses %>% 
  group_by(community_belonging, neighbourhood_rating) %>% 
  mutate(total_respondents = sum(n())) %>% 
  distinct(total_respondents) %>% 
  filter(neighbourhood_rating != "No opinion" &
           community_belonging != "Don't know") %>% 
  mutate(community_belonging = factor(community_belonging, 
                                      ordered = TRUE,
                                      levels = c("Very strongly",
                                                 "Fairly strongly",
                                                 "Not very strongly",
                                                 "Not at all strongly"))) %>% 
  mutate(neighbourhood_rating = factor(neighbourhood_rating, 
                                       ordered = TRUE,
                                       levels = c("Very good",
                                                  "Fairly good",
                                                  "Fairly poor",
                                                  "Very poor"
                                       ))) %>% 
  ggplot(aes(x = neighbourhood_rating, y = total_respondents, fill = community_belonging)) +
  geom_col(position = "dodge") +
  scale_fill_brewer(palette = "BrBG") +
  theme_minimal() +
  labs(title = "Neighbourhood rating by sense of community belonging",
       subtitle = "Scotland 2012-2019",
       x = "Neighbourhood rating",
       y = "Total respondents") +
  guides(fill = guide_legend(title = "Community belonging")) 
```

```{r}
neighbour_time <- shs_responses %>% 
  group_by(year, neighbourhood_rating) %>% 
  mutate(total_respondents = sum(n())) %>% 
  distinct(total_respondents) %>% 
  filter(neighbourhood_rating != "No opinion") %>% 
  mutate(neighbourhood_rating = factor(neighbourhood_rating, 
                                       ordered = TRUE,
                                       levels = c("Very good",
                                                  "Fairly good",
                                                  "Fairly poor",
                                                  "Very poor"
                                       ))) %>% 
  ggplot(aes(x = year, y = total_respondents, colour = neighbourhood_rating)) +
  geom_line() +
  scale_colour_brewer(palette = "BrBG") +
  theme_minimal() +
  labs(title = "Neighbourhood rating over time",
       subtitle = "2012-2019",
       x = "Year",
       y = "Total respondents") +
  theme(legend.title = element_blank())
```

```{r}
use_rating <- shs_responses %>% 
  group_by(nearest_green_space_use, neighbourhood_rating) %>% 
  mutate(total_respondents = sum(n())) %>% 
  distinct(total_respondents) %>% 
  filter(neighbourhood_rating != "No opinion") %>% 
  mutate(nearest_green_space_use = factor(nearest_green_space_use, 
                                          ordered = TRUE,
                                          levels = c("Every day",
                                                     "Several times a week",
                                                     "Once a week",
                                                     "Don't know",
                                                     "Several times a month",
                                                     "Once a month",
                                                     "Not at all"))) %>% 
  mutate(neighbourhood_rating = factor(neighbourhood_rating, 
                                       ordered = TRUE,
                                       levels = c("Very good",
                                                  "Fairly good",
                                                  "Fairly poor",
                                                  "Very poor"
                                       ))) %>% 
  ggplot(aes(x = neighbourhood_rating, y = total_respondents, fill = nearest_green_space_use)) +
  geom_col(position = "dodge") +
  scale_fill_brewer(palette = "BrBG") +
  theme_grey() +
  labs(title = "Neighbourhood rating by use of greenspace",
       subtitle = "Scotland 2012-2019",
       x = "Neighbourhood rating",
       y = "Total respondents") +
  guides(fill = guide_legend(title = "Use of greenspace")) +
  

```

```{r}
shs_responses %>% 
  count(nearest_green_space_use)
```

```{r}
shs_responses %>% 
  count(neighbourhood_rating)
```

```{r}
top_bottom_3 <- shs_data %>% 
  filter(feature_name %in% c("East Lothian", "Shetland Islands", "Dumfries And Galloway", "Glasgow City", "Dundee City", "West Dunbartonshire"))

top_bottom_plot <- top_bottom_3 %>% 
  filter(walking_distance_to_nearest_greenspace %in% c("More than 10 minutes", 
                                                       "Less than 10 minutes")) %>% 
  filter(measure == "Neighbourhood rating") %>% 
  filter(gender == "All" &
           urban_rural_classification == "All" &
           simd_quintiles == "All" &
           type_of_tenure  == "All" &
           household_type  == "All" &
           ethnicity == "All" &
           response != "No opinion") %>% 
  group_by(response, walking_distance_to_nearest_greenspace, feature_name) %>% 
  mutate(percentage_2013_2019 = mean(percent)) %>% 
  distinct(percentage_2013_2019) %>% 
  mutate(response = factor(response, 
                           ordered = TRUE,
                           levels = c("Very good",
                                      "Fairly good", 
                                      "Fairly poor",
                                      "Very poor"))) %>%
  mutate(walking_distance_to_nearest_greenspace = factor(walking_distance_to_nearest_greenspace)) %>% 
  mutate(feature_name = factor(feature_name,
                               ordered = TRUE,
                               levels = c("East Lothian", 
                                          "Shetland Islands", 
                                          "Dumfries And Galloway", 
                                          "Glasgow City", 
                                          "Dundee City", 
                                          "West Dunbartonshire"))) %>% 
  ggplot(aes(x = response, y = percentage_2013_2019)) +
  geom_col(position = "dodge", fill = "cyan4") +
  facet_wrap(~ feature_name) +
  #scale_fill_manual(values = c("darkslategray3", "deepskyblue4")) +
  theme_minimal() +
  labs(title = "Neighbourhood rating by area",
       subtitle = "Scotland 2013-2019",
       x = "Rating",
       y = "Percentage") +
  guides(fill = guide_legend(title = "Walking distance to nearest greenspace")) +
  coord_flip()

```

