---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---
# Scottish Household Survey EDA

# Read in the data ----

```{r}
library(tidyverse)
library(janitor)
library(here)
library(GGally)
library(readxl)
```

```{r}

community_belonging <- read_csv(here::here("raw_data/community_belonging.csv")) %>% 
  clean_names()

green_spaces <- read_csv(here::here("raw_data/green_spaces.csv")) %>% 
  clean_names()

neighbourhood_rating <- read_csv(here::here("raw_data/neighbourhood_rating.csv")) %>% 
  clean_names()

area_improvement <- read_csv(here::here("raw_data/area_improvement.csv")) %>% 
  clean_names()

neighbourhood_involvement <- read_csv(here::here("raw_data/neighbourhood_involvement.csv")) %>% 
  clean_names()

shs_responses <- read_csv(here::here("raw_data/shs_aggregate_responses.csv")) %>% 
  clean_names()

national_performance_framework <- read_xlsx(here::here("raw_data/npf_database.xlsx")) %>% 
  clean_names() # Not used in analysis
```
### Community Belonging data

```{r}
glimpse(community_belonging)
skimr::skim(community_belonging)
```
- No missing values
- 11 character vectors and 2 numerical (`date_code` aka year, `value` aka percentage)
- columns of particular interest for business question: 
  1. `community_belonging` (answers to the question "how strongly do you feel you belong to your immediate neighbourhood?")
  2. `walking_distance_to_nearest_greenspace` (could change to a numeric with units in minutes)
- are both `measurement` and `units` columns needed?
- what does `feature_code` relate to? Local authority? Something else?
- personal attributes: `gender`, `ethnicity`, `simd_quintiles`
- household attributes: `household_type`, `type_of_tenure`, `urban_rural_classification`
- includes aggregated data ("All")

```{r}
community_belonging %>% 
  count(feature_code)

community_belonging %>% 
  count(gender)

community_belonging %>% 
  count(ethnicity)

community_belonging %>% 
  count(walking_distance_to_nearest_greenspace)

community_belonging %>% 
  count(household_type)

community_belonging %>% 
  count(type_of_tenure)

```


```{r}
community_belonging %>% 
  count(community_belonging)

community_belonging %>% 
  ggplot() +
  geom_histogram(aes(value)) +
  facet_wrap(~ community_belonging)
```


### Greenspace data

```{r}
glimpse(green_spaces)
skimr::skim(green_spaces)
```

- same format as community_belonging data (partial match?)
- added `age` column (grouped)
- `distance__to_nearest_green_or_blue_space` is now measured in percentage

```{r}
green_spaces %>% 
  count(distance_to_nearest_green_or_blue_space)

green_spaces %>% 
  ggplot() +
  geom_histogram(aes(value)) +
  facet_wrap(~ distance_to_nearest_green_or_blue_space)

green_spaces %>% 
  filter(value == 0)
```


### Neighbourhood rating

```{r}
glimpse(neighbourhood_rating)
skimr::skim(neighbourhood_rating)
```

- again, same format
- added `neighbourhood_rating` column 

```{r}
# Trying to figure out what the measurement column pertains to
neighbourhood_rating %>% 
  filter(feature_code == "S12000005" &
         date_code == 2013 &
         gender == "All" &
         urban_rural_classification == "All" &
         simd_quintiles == "All" &
         type_of_tenure == "All" &
         household_type == "All" &
         ethnicity == "All" &
         walking_distance_to_nearest_greenspace == "All") %>% 
  group_by(measurement, neighbourhood_rating) %>% 
  mutate(total_value = sum(value)) %>% 
  select(value, total_value)
```

```{r}
neighbourhood_rating %>% 
  count(walking_distance_to_nearest_greenspace)

neighbourhood_rating %>% 
  count(neighbourhood_rating)

neighbourhood_rating %>% 
  ggplot() +
  geom_histogram(aes(value)) +
  facet_wrap(~ neighbourhood_rating)

neighbourhood_rating %>% 
  filter(value == 0)
```

### Area Improvement

```{r}
glimpse(area_improvement) 
skimr::skim(area_improvement)
```
- has additional `feature_name` and `feature_type` columns
- answers question whether selected adult thinks their area has improved

```{r}
area_improvement %>% 
  count(local_neighbourhood_improvement) # what do differences in response nos mean? surely they should be recorded for all?
```


```{r}
area_improvement %>% 
  count(feature_type)

area_improvement %>% 
  count(feature_name)
```

```{r}
area_improvement %>% 
  ggplot() +
  geom_histogram(aes(value)) +
  facet_wrap(~ local_neighbourhood_improvement)
```
```{r}
area_improvement %>% 
 filter(value == 0 & measurement == "Percent")
```



### Neighbourhood Involvement

```{r}
glimpse(neighbourhood_involvement)
skimr::skim(neighbourhood_involvement)
```

- answers question to do with involvement with people in neighbourhood (how does this interact with neighbourhood_rating or community_belonging?)

```{r}
neighbourhood_involvement %>% 
  count(involvement_with_people_in_the_neighbourhood)
```

- there's no negative framing of this question 

```{r}
neighbourhood_involvement %>% 
  ggplot() +
  geom_histogram(aes(value)) +
  facet_wrap(~ involvement_with_people_in_the_neighbourhood)
```



### SHS aggregated responses

```{r}
glimpse(shs_responses)
skimr::skim(shs_responses)
```

- quite a lot of missing values in `satisfaction_with_nearest_green_space`, `highest_education_level`, `nearest_green_space_use`, `volunteering_last_twelve_months` (and 1 in `economic_status`)
- 1 most common `n_persons` value
- age is categorical/factor



```{r}
shs_responses %>% 
  count(n_persons)

shs_responses %>% 
  count(satisfaction_with_nearest_green_space)

shs_responses %>% 
  count(highest_education_level)

shs_responses %>% 
  count(volunteering_last_twelve_months)

shs_responses %>% 
  count(gender)

shs_responses %>% 
  count(age)
```

```{r}
shs_responses %>% 
  filter(is.na(satisfaction_with_nearest_green_space) & is.na(nearest_green_space_use)) %>% 
  count(year)

shs_responses %>% 
  count(year)

shs_responses %>% 
  group_by(year) %>% 
  summarise(total = n()) %>% 
  summarise(mean(total))
```
- looks like greenspace questions were not asked in 2017 or 2018 (NMAR)

```{r}
shs_responses %>% 
  filter(is.na(volunteering_last_twelve_months)) %>% 
  count(year)
```
- volunteering was left out of 2012 survey (NMAR)



```{r}
shs_responses %>% 
  filter(is.na(highest_education_level)) %>% 
  count(year)
```

- `highest_education_level` has more of an unstructered nature to NAs (MAR)

```{r}
shs_responses %>% 
  filter(is.na(economic_status))
```


```{r}
shs_responses %>% 
  filter(n_persons > 1) %>% 
  count(year)
```

- if limiting data to responses of individuals lose high proportion of data from 2012/17/18 
- 2012 would be 80% complete
- 2017 and 2018 would be roughly 69% complete
- Is it worth it? 

```{r}
shs_responses %>% 
  distinct(nearest_green_space_use)

shs_responses %>% 
  distinct(volunteering_last_twelve_months)
```

```{r}
shs_responses %>% 
  ggplot() +
  geom_bar(aes(x = neighbourhood_rating))
```



```{r, message=FALSE}

split1 <- shs_responses_imp %>% 
  select(neighbourhood_rating, 
         community_belonging, 
         volunteering_last_twelve_months)

split2 <- shs_responses_imp %>% 
  select(neighbourhood_rating, 
         year, 
         age, 
         economic_status,
         household_size,
         highest_education_level)

split3 <- shs_responses_imp %>% 
  select(neighbourhood_rating,
         distance_to_nearest_green_space,
         satisfaction_with_nearest_green_space,
         nearest_green_space_use)

ggpairs(split1)
ggpairs(split2)
ggpairs(split3)

```


### National Performance Framework

```{r}
glimpse(national_performance_framework)
skimr::skim(national_performance_framework %>% 
              filter(source == "Scottish Household Survey"))
```

- no missing values
- should only select data relevant to task

```{r}
national_performance_framework %>% 
  filter(source == "Scottish Household Survey") %>% 
  count(measure)
```

```{r}
national_performance_framework %>% 
  filter(str_detect(indicator, "green") & 
         outcome == "Communities")  


national_performance_framework %>% 
  filter(str_detect(indicator, "green") & 
         outcome == "Communities" &
         characteristic == "Age") %>% 
  mutate(year = as.numeric(year),
         figure = as.numeric(figure)) %>% 
  ggplot(aes(x = year, y = figure, colour = breakdown)) +
  geom_line() 
```


```{r}
national_performance_framework %>% 
  filter(str_detect(indicator, "green") & year == 2020)
```
- the 2020 question looks like it was intended to inform environmental policy not communities

```{r}
npf <- read_csv(here::here("clean_data/npf_shs_communities.csv"))
```

```{r}
summary(npf)
glimpse(npf)
```

```{r}
community_belonging <- read_csv(here::here("clean_data/community_belonging_clean.csv"))
glimpse(community_belonging)
```

```{r}
community_belonging %>% 
  pivot_longer(community_belonging,
               names_to = "measure",
               values_to = "response") %>% 
  relocate(measure, .after = date_code) %>% 
  relocate(response, .before = percent)
```



