---
title: "R Notebook"
output: html_notebook
---
# Scottish Household Survey EDA

Read in the data 

```{r}
library(tidyverse)
library(janitor)
```

```{r}
community_belonging <- read_csv("raw_data/community_belonging.csv") %>% 
  clean_names()

green_spaces <- read_csv("raw_data/green_spaces.csv") %>% 
  clean_names()

neighbourhood_rating <- read_csv("raw_data/neighbourhood_rating.csv") %>% 
  clean_names()
```
### Community Belonging data

```{r}
glimpse(community_belonging)
skimr::skim(community_belonging)
```
> No missing values
> 11 character vectors and 2 numerical (`date_code` aka year, `value` aka percentage)
> columns of particular interest for business question: 
  1. `community_belonging` (answers to the question "how strongly do you feel you belong to your immediate neighbourhood?")
  2. `walking_distance_to_nearest_greenspace` (could change to a numeric with units in minutes)
> are both `measurement` and `units` columns needed?
> what does `feature_code` relate to? Local authority? Something else?
> personal attributes: `gender`, `ethnicity`, `simd_quintiles`
> household attributes: `household_type`, `type_of_tenure`, `urban_rural_classification`
> includes aggregated data ("All")

```{r}
community_belonging %>% 
  count(feature_code)

community_belonging %>% 
  count(gender)

community_belonging %>% 
  count(ethnicity)

community_belonging %>% 
  count(walking_distance_to_nearest_greenspace)

community_belonging %>% 
  count(household_type)

community_belonging %>% 
  count(type_of_tenure)

community_belonging %>% 
  count(community_belonging)
```

### Greenspace data

```{r}
glimpse(green_spaces)
skimr::skim(green_spaces)
```

> same format as community_belonging data (partial match?)
> added `age` column (grouped)
> `distance__to_nearest_green_or_blue_space` is now measured in percentage

```{r}
green_spaces %>% 
  count(distance_to_nearest_green_or_blue_space)
```


### Neighbourhood rating

```{r}
glimpse(neighbourhood_rating)
skimr::skim(neighbourhood_rating)
```

> again, same format
> added `neighbourhood_rating` column 

```{r}
# Trying to figure out what the measurement column pertains to
neighbourhood_rating %>% 
  filter(feature_code == "S12000005" &
         date_code == 2013 &
         gender == "All" &
         urban_rural_classification == "All" &
         simd_quintiles == "All" &
         type_of_tenure == "All" &
         household_type == "All" &
         ethnicity == "All" &
         walking_distance_to_nearest_greenspace == "All") %>% 
  group_by(measurement, neighbourhood_rating) %>% 
  mutate(total_value = sum(value)) %>% 
  select(value, total_value)
```

```{r}
neighbourhood_rating %>% 
  count(walking_distance_to_nearest_greenspace)

neighbourhood_rating %>% 
  count(neighbourhood_rating)
```

