---
title: "Random Forest Model"
output: html_notebook
---
# Load in libraries ----
```{r}
library(janitor)
library(here)
library(rpart)
library(rpart.plot)
library(ranger)
library(yardstick)
library(caret)
library(tidyverse)
```

# Read in data ----

```{r}
shs_responses <- read_csv(here::here("clean_data/shs_responses_clean.csv"))

glimpse(shs_responses)

```

# Data Cleaning ----

```{r}
shs_responses <- shs_responses %>% 
  mutate(across(where(is_character), as_factor)) 
 

# Order factors
shs_responses$neighbourhood_rating <- ordered(shs_responses$neighbourhood_rating,
                                              levels = c("Very good", 
                                                         "Fairly good", 
                                                         "No opinion", # middle position as proxy for neutrality
                                                         "Fairly poor",
                                                         "Very poor"))

shs_responses$community_belonging <- ordered(shs_responses$community_belonging,
                                             levels = c("Very strongly",
                                                        "Fairly strongly",
                                                        "Don't know",
                                                        "Not very strongly",
                                                        "Not at all strongly"))

shs_responses$distance_to_nearest_green_space <- ordered(shs_responses$distance_to_nearest_green_space,
                                                         levels = c("Less than 5 minutes",
                                                                    "Between 6 and 10 minutes",
                                                                    "Between 11 and 20 minutes",
                                                                    "Between 21 and 30 minutes",
                                                                    "More than 30 minutes",
                                                                    "Don't know"))
shs_responses$satisfaction_with_nearest_green_space <- ordered(shs_responses$satisfaction_with_nearest_green_space,
                                                               levels = c("Very satisfied",
                                                                          "Fairly satisfied",
                                                                          "No opinion",
                                                                          "Fairly dissatisfied",
                                                                          "Very dissatisfied"))

shs_responses$age <- ordered(shs_responses$age,
                             levels = c("16 - 34 Years",
                                        "35 - 64 Years",
                                        "65 + Years"))

shs_responses$nearest_green_space_use <- ordered(shs_responses$nearest_green_space_use,
                                                 levels = c("Every day",
                                                            "Several times a week",
                                                            "Once a week",
                                                            "Don't know",
                                                            "Several times a month",
                                                            "Once a month",
                                                            "Not at all"))

skimr::skim(shs_responses)
```

## Test/train/validate split

```{r}
test <- slice_sample(shs_responses, prop = 0.2)
validate <- slice_sample(shs_responses, prop = 0.2)
train <- slice_sample(shs_responses, prop = 0.6) 
```


# Random Forest Model ----

```{r}
# Create RF model
rf_classifier <- ranger(neighbourhood_rating ~.,
                        data = train, 
                        importance = "impurity",
                        num.trees = 1000, # num of trees to construct
                        mtry = 2, # num of variables to split at each node
                        min.node.size = 5) # regression model

rf_classifier
importance(rf_classifier)
sqrt(rf_classifier$prediction.error) # OOB RMSE
```
OOB 39%
OOB RMSE 0.625

_Most important variables_
1. `community_belonging` 
2. `year`
3. `satisfaction_with_nearest_green_space`

_Least important variables_
1. `volunteering_last_twelve_months`
2. `gender`
3. `age`

## Test model

```{r}
rf_test_pred <- test %>% 
  mutate(pred = predict(rf_classifier, data = test)$predictions)

confusionMatrix(rf_test_pred$pred, rf_test_pred$neighbourhood_rating)
```
Accuracy: 68%
Better sens/spec for Very good and Fairly good (to do with the data?) 

## Validate model

```{r}
rf_validate_pred<- validate%>% 
  mutate(pred = predict(rf_classifier, data = validate)$predictions)

rf_validate_pred %>% 
  select(neighbourhood_rating, pred)

confusionMatrix(rf_validate_pred$pred, rf_validate_pred$neighbourhood_rating)
```


# Fine-tuning the model ----

```{r}
# change the number of variables at each node
# Boehmke & Greenwell (2020) suggest floor(num of features / 3)

n_features <- length(setdiff(names(shs_responses), "neighbourhood_rating"))

rf_classifier_2 <- ranger(neighbourhood_rating ~.,
                        data = train, 
                        importance = "impurity",
                        num.trees = 1000, 
                        mtry = floor(n_features/3), 
                        min.node.size = 5)

rf_classifier_2
sqrt(rf_classifier_2$prediction.error) 
importance(rf_classifier_2)
```
OOB 39.5%
OOB RMSE 0.628

## Test the model
```{r}
rf_test_pred_2<- test %>% 
  mutate(pred = predict(rf_classifier_2, data = test)$predictions)

confusionMatrix(rf_test_pred_2$pred, rf_test_pred_2$neighbourhood_rating)
```

Accuracy = 75%
sens/spec scores are better all-round


```{r}
# Try a third model with more trees

rf_classifier_3 <- ranger(neighbourhood_rating ~.,
                        data = train, 
                        importance = "impurity",
                        num.trees = 10000, # 10000 trees
                        mtry = 2, 
                        min.node.size = 5) 

rf_classifier_3
importance(rf_classifier_3)
sqrt(rf_classifier_3$prediction.error) 
```
OOB 38.9%
OOB RMSE 0.624

```{r}
rf_test_pred_3<- test %>% 
  mutate(pred = predict(rf_classifier_3, data = test)$predictions)

confusionMatrix(rf_test_pred_3$pred, rf_test_pred_3$neighbourhood_rating)
```
Accuracy 68% 
Previous model better

## Validate model 
The second model is the preferred

```{r}
rf_validate_pred_2<- validate%>% 
  mutate(pred = predict(rf_classifier_2, data = validate)$predictions)

rf_validate_pred_2 %>% 
  select(neighbourhood_rating, pred)

confusionMatrix(rf_validate_pred_2$pred, rf_validate_pred_2$neighbourhood_rating)
```


```{r}
# Plot the importance of variables

plot <- vip::vip(rf_classifier_2, num_features = 11, aesthetics = list(fill = "deepskyblue4"))

plot
```
